Research Notes

### 2018-08-14 writeups/notes.tex

If we approximate $F(x, z, t) \approx F_0\delta\p*{z - z_0} e^{i(k_xx - \omega t)}$
for the time being, then clearly the only term that can be equally singular is
the $\ptd{}{z}$ term, leaving us with
\begin{equation}
    -\omega^2 \ptd{u_{1z}}{z} - \frac{Fgk_x^2}{\rho_0} = 0.
\end{equation}
Since $u_{1z}$ both above and below $z_0$ satisfy the homogeneous PDE, which has
plane wave solutions with an $e^{z/2H}$ envelope where $k_zH \gg 1$, we can
integrate about a small domain around $z_0$ and obtain
\begin{equation}
    \at{\pd{u_{1z}}{z}}_{z > z_0} - \at{\pd{u_{1z}}{z}}_{z < z_0}
        = -F_0\frac{gk_x^2}{\rho_0\omega^2}e^{i(k_xx - \omega t)}.
\end{equation}
It should then be apparent that the $u_{1z}$ solution is
\begin{equation}
    u_{1z} = \frac{F_0gk_x^2}{\rho_0(z = z_0)\omega^2}\frac{1}{2ik_z} \times
    \begin{cases}
        -e^{(z - z_0)/2H}e^{i(k_xx + k_z(z - z_0) - \omega t)} & z > z_0\\
        e^{(z - z_0)/2H}e^{i(k_xx + k_z(z - z_0) - \omega t)} & z < z_0\\
    \end{cases}.
\end{equation}

@@

>>>>

If we approximate $F(x, z, t) \approx F_0\delta\p*{z - z_0} e^{i(k_xx - \omega t)}$
for the time being, then clearly the only term that can be equally singular is
the $\ptd{}{z}$ term, leaving us with
\begin{equation}
    -\omega^2 \ptd{u_{1z}}{z} - \frac{Fgk_x^2}{\rho_0} = 0.
\end{equation}
Since $u_{1z}$ both above and below $z_0$ satisfy the homogeneous PDE, which has
plane wave solutions with an $e^{z/2H}$ envelope where $k_zH \gg 1$, we can
integrate about a small domain around $z_0$ and obtain
\begin{equation}
    \at{\pd{u_{1z}}{z}}_{z > z_0} - \at{\pd{u_{1z}}{z}}_{z < z_0}
        = -F_0\frac{gk_x^2}{\rho_0\omega^2}e^{i(k_xx - \omega t)}.
\end{equation}
It should then be apparent that the $u_{1z}$ solution is
\begin{equation}
    u_{1z} = \frac{F_0gk_x^2}{\rho_0(z = z_0)\omega^2}\frac{1}{2ik_z} \times
    \begin{cases}
        e^{(z - z_0)/2H}e^{i(k_xx + k_z(z - z_0) - \omega t)} & z > z_0\\
        e^{(z - z_0)/2H}e^{i(k_xx - k_z(z - z_0) - \omega t)} & z < z_0\\
    \end{cases}.
\end{equation}

@@

\chapter{Stratified Flow}

\section{Linear Breaking Criterion}

Since we're having numerical difficulties, we will simultaneously consider
\emph{corotation resonances} as a pathway to instability. This corresponds in
the 2D case to a shear flow $\vec{u}_0 = u_0(z)\hat{x}$. We can do the math for
this out, but the end result produces dispersion relation
\begin{equation}
    \p*{\omega - u_0k_x}^2 = \frac{k_x^2}{k_x^2 + k_z^2 + \frac{1}{4H^2}}
        \p*{N^2 - \omega \pd{u_0}{z}}.
\end{equation}
Since our fluid is generally not stratified strongly and $\omega \ll N$ in the
IGW regime, we can drop the $\pd{u_0}{z}$ term and the shear flow becomes a
simple Doppler shift of $\omega$. Denote $\omega_{cr}(z) = \omega - u_0(z)k_x$
the corotating frequency. A singularity is obvious at $\omega_{cr} = 0$, exactly
the \emph{corotation resonance}, and one expects that as
\begin{equation}
    k_z(z) \approx \frac{k_xN}{\omega_{cr}(z)} \sim \frac{\omega_{cr}}{u_{1z}},
        \label{eq:corot_nonlin_cond}
\end{equation}
that nonlinear effects should kick in.

One advantage of a shear flow is that, instead of requiring $2H$ of vertical
height for amplitude growth, we can just choose $u_0 =
\frac{\omega}{k_x}\frac{z}{z_{\max}}$ to meet the corotation citerion at $z =
z_{\max}$. As $\pd{u_0}{z} \sim \frac{\omega}{k_xz_{\max}} \approx
\frac{N}{k_zz_{\max}}$, to meet the $N^2 \gg \omega \pd{u_0}{z}$ criterion we
require $N \gg \frac{\omega}{k_zz_{\max}}$ which is always true. We can even
induce the corotation resonance well below $z_{\max}$ and still be perfectly
within the regime of validity of our assumption.

A second advantage of a shear flow is that we can force at very small amplitudes
and still observe resonances, since the resonance is driven by the convergence
of $\omega_{cr} \to 0$ (purely coordinate driven) instead of $u_{1z} \sim
v_{ph}$ (dependent on wave amplitude).

The width of the dissipation zone can be estimated by noticing $u_0k_x = \omega
- \sqrt{k_xNu_{1z}}$ from the nonlinearity criterion
\autoref{eq:corot_nonlin_cond}. Thus, the width is roughly of order
$2\sqrt{k_xNu_{1z}}$, which has the right scaling behavior (a stronger
perturbation should be resonant farther away). Note however that, since $\omega$
is small, this width can easily be large compared to $\omega$!
